---
title: "Figure 1"
output: html_document
date: '2022-10-06'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prepare the libraries required for plotting
```{r library, echo=FALSE}
library(ggplot2)
library(tidyverse)
library(cowplot)
library(colorspace)
library(ggrepel)
library(viridis)
library(ggpubr)
library(dplyr)
library(ggpubr)
```
Figure 1: Generate diagram to introduce the dataset information used in the comparative study 
```{r diagram, Figure 1,echo = FALSE}
sample_info = data.frame(Cell_type =c('Naive B',"Naive B","HSC","HSC","Immature B","LargeB","Simulation *4","Technical"), Technology = c("FACS","Droplet","FACS","Droplet","FACS","Droplet","Droplet",'CEL-seq'), Size = c(1166,49,1174,486,44,10085,400,1015),Type = c(rep('Biological samples',6),"Simulation samples","Technical samples"), Method = c("full-length (smartseq2)","3'-end  (10X droplet)","full-length (smartseq2)","3'-end  (10X droplet)","full-length (smartseq2)","3'-end  (10X droplet)","3'-end  (10X droplet)","3'-end  (10X droplet)") )
sample_info <- sample_info[order(sample_info$Size,decreasing = T),]

region_cols <- c("#F0E442", "#0072B2", "#999999")
 
g1 <- ggplot(sample_info, aes(Cell_type  ,Method)) +
  geom_point(
    aes(color = Type, fill = Type,size =  Size )
    ,shape=16
  ) +scale_size(range = c(5, 12))+
   scale_color_viridis(discrete = TRUE, option = "D")+
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal_hgrid(12, rel_small = 1)+xlab('')+ylab('') # font size 12 pt throughout
ggsave("revised_figure/Dataset_info.png", width = 12, height = 6)
 
```
 
Figure 2: The evaluation of the metric performance with two main sequencing platforms

```{r, Figure 2 }
# load the final performance matrix 
Performance_on_all_gene <- readRDS(".~/Performance_on_all_genes.rds")
# reorder 
Performance_on_all_gene$metric <- factor(Performance_on_all_gene$metric, levels = c("SD", "IQR", "MAD", "CV","FF","LCV","BASiCS","DESeq2","DM","edgeR","glmGamPoi","scran","Seurat_mvp","Seurat_vst"))
 
as_tibble(Performance_on_all_gene) %>% 
    filter(cell_type%in%c("HSC","Naive B"))   %>% 
 ggplot(., aes(x=as.numeric(value), y=cell_type, fill=platform)) + 
  geom_boxplot(outlier.size = 0.1)+
  theme_bw(base_size = 12 )+
  facet_wrap(  .~metric ,nrow = 3,ncol=5,scales = "free_x" , dir="h")  +
  labs(title = "Cell-to-cell variability for all genes",
       y = "", x = "")  + theme(legend.position="bottom")
ggsave('Figure2a.png', width = 12, height =8 )

disatnce_table <-  Distance_between_cell_types
rownames(disatnce_table) <- c("HSC","full-length (smartseq2)","Naive B","3'-end (10X droplet)")
disatnce_table1 <- reshape2::melt(disatnce_table)
disatnce_table1$info <- rep(c("Platform variation","Cell type variation","Platform variation","Cell type variation"), 14)
disatnce_table1$Var1 <- as.factor(disatnce_table1$Var1)
disatnce_table1$Var1 <- factor(disatnce_table1$Var1,levels = c("HSC","Naive B","full-length (smartseq2)","3'-end (10X droplet)"))
library(ggplot2)
ggplot(data=disatnce_table1, aes(x=value, y=Var1, fill=info)) +scale_fill_manual(values=c('#d7aeac',"#5c6c7c"))+
  geom_bar(stat="identity", width=0.8) +
  theme_bw(base_size = 12 )+
  facet_wrap(  .~Var2 ,nrow = 3,ncol=5,  dir="h") +
  labs(title = "Distances between samples and sequencing platforms",
       y = "", x = "")  + theme(legend.position="bottom")
ggsave('Figure2b.png', width = 12, height =8 )
 
```

Figure 3: The evaluation of the metric comparison on different sample sizes (TMS and external)

```{r, Figure 3}
Performance_sample_size <- Performance_on_all_gene[Performance_on_all_gene$platform=="full-length (smartseq2)",]

fig3a1 <- ggplot(Performance_sample_size, aes(x = as.numeric(value), colour = cell_type)) +
  geom_density(lwd = 1.2, linetype = 1, adjust = 10) + 
  scale_color_manual(values = c("#F76D5E", "#72D8FF", "#01524a"))+facet_wrap(.~metric,nrow = 3,ncol = 5, scales='free')+xlab("Cell-to-cell variability (full-length (smartseq2))") +theme_bw(base_size = 12)
 
Performance_sample_size1 <- Performance_on_all_gene[Performance_on_all_gene$platform!="full-length (smartseq2)",]
levels(Performance_sample_size1$cell_type)[4] <- 'B cell'

fig3a2 <- ggplot(Performance_sample_size1[Performance_sample_size1$cell_type%in%c("Naive B","B cell"),], aes(x = as.numeric(value), colour = cell_type)) +
  geom_density(lwd = 1.2, linetype = 1, adjust = 50) + 
  scale_color_manual(values = c("#72D8FF","#6699CC" ))+facet_wrap(.~metric,nrow = 3,ncol = 5, scales='free')+xlab("Cell-to-cell variability (3'-end (10X droplet))") +theme_bw(base_size = 12)  

ggarrange(fig3a1+ggtitle("Cell-to-cell variability (full-length (smartseq2))"),fig3a2+ggtitle("Cell-to-cell variability (3'-end (10X droplet))"), ncol = 1, nrow = 2,labels = 'auto',widths  = c(2,1))

ggsave("Figure3.png", width = 10, height = 10)
```

Figure 4: 
a) The evaluation of the metric comparison on known-degree of variability 
b) The overall performance score plotted as a dotplot

```{r, Figure 4}
Performance_lowvar_genes <- readRDS('D:/PhD/Trip to China/revised scripts/Performance_on_lowvar_genes.rds')
Performance_lowvar_genes.1 <- reshape2::melt(Performance_lowvar_genes) 
Performance_sim <- readRDS('D:/PhD/Trip to China/revised scripts/Performance_on_simulation.rds')
Performance_sim <- data.frame(metric=Performance_sim$variable,variable=rep('Simulation (HVGs)',48),value = Performance_sim$value/200)

Performance_known_var <- rbind(Performance_lowvar_genes.1[,c('metric','variable','value')],Performance_sim)
col <- c("#70AD47","#70AD47","#70AD47","#70AD47", "#70AD47",'#ff0000',"#FFC000","#FFC000","#FFC000","#FFC000","#FFC000","#FFC000","#FFC000","#000000")#outlier.size = 0.6
p_lowvar <- ggplot(Performance_known_var, aes(x=metric, y=as.numeric(value), fill=variable)) + 
  geom_boxplot() +
  theme_bw(base_size = 12) + scale_fill_brewer(palette="Dark2")  +
  labs(title = "Rediscovery rate in the presence of known degrees of variability",
       y = "", x = "") + theme(axis.text = element_text(size = 12),axis.text.x = element_text(colour = col ))  + labs(fill='') 
# Metric performance 
Ranks <- readRDS('D:/PhD/Trip to China/revised scripts/Final_ranks.rds')
Seperate_Ranks <- readRDS('D:/PhD/Trip to China/revised scripts/Overall_evaluation.rds')


p1 <- ggplot(Ranks, 
       aes(x= rank , 
           y=reorder(metric, rank))) +
  geom_point()+theme_bw()+ylab('')  +
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank() 
  )
ggsave('overall_ranks.png',height = 16, width = 4) 

a <- c("#70AD47","#70AD47","#70AD47","#FFC000", '#ff0000' ,"#FFC000","#FFC000","#FFC000","#70AD47","#FFC000","#70AD47","#FFC000","#000000","#FFC000")

Seperate_Ranks$variable <- factor(Seperate_Ranks$variable,levels=c("Zero","Mean","Ribosomal","SEGs",'Gene Length'))
Seperate_Ranks$platform <- as.factor(Seperate_Ranks$platform)
levels(Seperate_Ranks$platform) <- c("3'-end (10X droplet)","full-length (smartseq2)")
p <- ggplot(Seperate_Ranks , aes(x =  value , y = metric))    +
  geom_point(aes(colour = platform, size=size), alpha=0.8) +facet_grid(cols = vars(variable))+theme_bw() +
  theme(axis.text.y = element_text( hjust = 1, colour = a,size=12))+xlab("Performance")+ylab('')+ 
      theme(strip.text.x = element_text(size = 14 ))
ggarrange(p_lowvar,p,ncol = 1, nrow = 2, labels = 'auto',heights = c(1.5,2))
ggarrange(p_lowvar, ncol = 1, nrow = 1, labels = 'auto' )
ggarrange(p , ncol = 1, nrow = 1, labels = 'b' )

ggsave("Figure4.png", width =20, height = 16)

Overall_smart <- Overall[Overall$platform=="full-length (smartseq2)",]
Overall_smart$metric <- factor(Overall_smart$metric, levels=levels(Ranks_smart$metric)) # match with ranks
Overall_smart$type <- unname(sapply(Overall_smart$type, function(x)stringr::str_split(x,'_')[[1]][1]))

p_smart <- ggplot( Overall_smart, aes(x = (1-value), y = metric))+  geom_point(aes(colour = type, size=size ), alpha=0.8) +facet_grid(cols = vars(variable))+theme_bw() +
  theme(axis.text.y = element_text( hjust = 1, colour = a,size=14))+xlab("Performance")+ylab('') 


Overall_10x <- Overall[Overall$platform!="full-length (smartseq2)",]
Overall_10x$metric <- factor(Overall_10x$metric, levels=levels(Ranks_10x$metric)) # match with ranks 
Overall_10x$type <- unname(sapply(Overall_10x$type, function(x)stringr::str_split(x,'_')[[1]][1]))

b <- c("#70AD47","#70AD47","#70AD47","#FFC000", '#ff0000' ,"#FFC000","#FFC000","#FFC000","#70AD47","#FFC000","#FFC000","#FFC000","#000000","#70AD47")
p_10x <- ggplot(Overall_10x , aes(x = (1-value), y = metric))    +
  geom_point(aes(colour = type, size=size), alpha=0.8) +facet_grid(cols = vars(variable))+theme_bw() +
  theme(axis.text.y = element_text( hjust = 1, colour = b,size=14),legend.position = 'bottom')+xlab("Performance")+ylab('')



ggarrange(p , p1,legend = 'right', widths = c(8,1))
ggarrange(p_smart + theme(legend.position="bottom") , p1_smart  ,widths = c(8,1), heights = c(1.3,1))
ggsave('ranks_smart.png',width = 16, height = 6)
ggarrange(p_10x+ scale_color_manual(values=c("#F8766D","#7CAE00","#00BFC4" )) , p1_10x  ,widths = c(8,1))
ggsave('ranks_10x.svg',width = 16, height = 6)
```


```{r supp. }
Seperate_Ranks <- readRDS('D:/PhD/Trip to China/revised scripts/Overall_evaluation.rds')

Ranks_smart <- aggregate(Overall_smart$value, by=list(as.character(Overall_smart$metric)),FUN=mean)
colnames(Ranks_smart) <- c('metric',"performance")
Ranks_smart <- Ranks_smart[order(Ranks_smart$performance,decreasing = T),]
Ranks_smart$metric <- factor(Ranks_smart$metric, levels = Ranks_smart$metric)

Ranks_10x <- aggregate(Overall_10x$value, by=list(as.character(Overall_10x$metric)),FUN=mean)
colnames(Ranks_10x) <- c('metric',"performance")
Ranks_10x <- Ranks_10x[order(Ranks_10x$performance,decreasing = T),]
Ranks_10x$metric <- factor(Ranks_10x$metric, levels = Ranks_10x$metric)

```

 